---
title: "Practical"
author: "K MWai"
output: beamer_presentation
header-includes:  
- \usetheme{CambridgeUS}
- \useinnertheme{rounded}
- \usecolortheme{dolphin}
- \usefonttheme{professionalfonts}
- \usepackage{xcolor,soul,hyperref,cite,graphicx,amsmath}
linkcolor: blue
urlcolor: blue
---

\begin{center}

\textbf{Measures of Data}\\      


\end{center}

## Exercise. Q1 

The lengths of time, in minutes, that 10 patients waited in a doctor's office before receiving treatment
were recorded as follows: 5, 11, 9, 5, 10, 15, 6, 10, 5, and 10. Treating the data as a random sample.

Find    
(a) the mean;     
(b) the median;     
(c) the mode;    
(d) the range;     
(e) the standard deviation.


## Exercise. Q2

In the 2014-15 rugby season, Kemri University had the following score differences for the 13 games it played: 11 49 32 3 6 38 38 30 8 40 31 5 36.     

Find      
(a) the mean score difference;        
(b) the median score difference.       

## Solutions. Q1
\small
```{r,echo=T,comment=NA}
time<-c(5, 11, 9, 5, 10, 15, 6, 10, 5,10)
# a) mean
mean(time)
# b) median
median(time)
# c) mode
sort(table(time))

```
***

## Solutions. Q1
\small
```{r,echo=T,comment=NA}
# d) range
range(time)
# e) sd
sd(time)

```

## Solutions. Q2
\small
```{r,echo=T,comment=NA}
rugby<-c(11,49,32,3,6,38,38,30,8,40,31,5,36)
# a) mean
mean(rugby)
# b) median
median(rugby)
# c) mode
sort(table(rugby))


```
## Solutions. Q2
\small
```{r,echo=T,comment=NA}
# d) range
range(rugby)
# e) sd
sd(rugby)

```


* * *

Lets go Plotting 




\begin{center}

\textbf{Graphical Exploration}\\      


\end{center}


## Exercise. 1

Draw a box plot to show the distribution of the birth weight(bweight) for the different age groups (agegrp).        

- Label the titles i.e y, x, and the main title      

- Change the color of the titles from the default black to a color of your choice.

- Increase the size of axis labels by 20%, main label by 35%

- y scale to start from 500 to 5000 grams         

- Get the median birthweight for the age categories      

- save the graph as a .pdf file to a location of your choice        


## Exercise. 2

Draw a Grouped bar plot showing the number of kids in each category of lbw grouping them by sex.      


- Label the titles i.e y, x, and the main title      

- Change the color of the titles from the default black to a color of your choice.

- Increase the size of axis labels by 20%, main label by 35%

- Get the median birthweight for the age categories      

- save the graph as a .png file to a location of your choice    

## Solution 1

\small
```{r}
library(foreign)
birth<-read.csv("data/birthweight2.csv")
attach(birth)
```

```{r,echo=TRUE,fig.height=3.5,fig.width=6.8,results='hide'}
pdf("soln1.pdf")

boxplot(bweight~agegrp,las=1,main="Birth weight per age group",
    ylab="Birth weight(g)",ylim=c(500,5000),col.lab="darkgreen",
    col.main="red",xlab="Age group",cex.lab=1.2,cex.main=1.35)

dev.off()
```

## Solution 1 ... 
 
\begin{figure}[htbp!]
\centering
 \includegraphics[width=0.77\textwidth,height =0.85\textheight]{soln1}
\end{figure}

## 

\small

```{r,echo=T,comment=NA}

tapply(bweight,agegrp,median)



```

##  Solution 2

```{r, echo=F}
library(knitr)
tabs<-table(lbw,sex)
kable(tabs)
```


## Solution 2

\small
```{r,echo=TRUE,fig.height=3.3,fig.width=6.7,results='hide'}
pdf("soln2.png")

barplot(tabs,las=1,beside = T, main="Grouped Bar Plot",
        xlab="Sex",ylab="Counts", border="darkblue",
        col=c("darkgreen","lightgrey"),cex.lab=1.2,
        cex.main=1.35,col.lab="darkgreen",col.main="red")
legend("center",legend=(rownames(tabs)),
       fill=c("darkgreen","lightgrey"))
box()

dev.off()
```

## Solution 2 ... 
 
\begin{figure}[htbp!]
\centering
 \includegraphics[width=0.81\textwidth,height =0.85\textheight]{soln2}
\end{figure}




