---
title: "Graphical Exploration"
author: "K.Mwai"
output: beamer_presentation
header-includes:  
- \usetheme{CambridgeUS}
- \useinnertheme{rounded}
- \usecolortheme{dolphin}
- \usefonttheme{professionalfonts}
- \usepackage{xcolor,soul,hyperref,cite,graphicx,amsmath,verbatim,longtable}
linkcolor: blue
urlcolor: blue
---

#### Working Directroy
? is it set

## Graphical exploration of data.    
\normalsize
\begin{itemize}
\item [-] To have an understanding of you data we normally conduct exploratory data analysis (EDA) which can be graphical or numerical.  
\item []

\item [-] Primarily EDA is for seeing what the data can tell us before the formal modeling or hypothesis testing task.    
\item []
\item [-] Typical graphical techniques used in EDA are:    
\begin{itemize}
      \item [-.] Scatter plots,   
      \item [-.] Box plots,    
      \item [-.] Bar plots.
\end{itemize}
\end{itemize}


## Data

```{r,include=FALSE}
library(foreign)
birth<-read.csv("data/birthweight2.csv")
dim(birth)
```
\small
```{r,echo=F,results='asis'}
library(knitr)
kable(head(birth[,-7]),format = "markdown",padding = 0,longtable=T)
```

***

##      

\begin{center}

\textbf{Scatter plot}

\end{center}

## Scatter plot
\begin{itemize}
\item [-] Its a useful summary of a set of bivariate data (two variables) 
\item []
\item [-] It pairs up values of two quantitative variables in a data set with the aim of giving a good visual picture of the relationship between the two variables.   
\item []
\item [-] The resulting pattern indicates the type and strength of the relationship between the two variables.
\item []
\item [-] Usually drawn before working out a linear correlation coefficient or fitting a regression line.  

\end{itemize}

## Parts of a graph
 
\begin{figure}[htbp!] 
\centering
\includegraphics[width=0.75\textwidth,height =0.85\textheight]{graphParts}
\end{figure}


## Graphical parameters
-. Help in customizing features of your graphs
\small
```{r, eval=FALSE}
# Text and symbol size.         
cex = ;change size relative to the (default=1),1.5 is 50% larger,
      0.5 is 50% smaller.      
      cex.axis,cex.lab,cex.main       
# Plotting symbols
pch = ;0:25, +,*, -,% (default=1, open circle)        
# Lines
lty = ;line type (default is solid line), can be dashed, dotted,...     
lwd = ;line width. 2 is twice as wide.           
# Colors
col = ;0:8, col= 1, col="white".    
       col.axis, col.main, col.lab, bg, fg    
# Axes 
xlab="X-axis label",main="title",ylab="y-axix label",
xlim=c(xmin, xmax)
```


## Scatter plot ... `R code`    
\small

```{r ,fig.height=5,fig.width=12}
plot(birth$gestwks,birth$bweight,las=1,xlab="Gestation(weeks)",
     ylab="Birth weight",col.lab="darkgreen", cex.lab=1.2,
     main="Weight at Birth",col.main="blue",cex.main=1.4,
     xlim=c(25,42),col="darkred",type="p", cex=0.9,pch=1)
```

## Scatter plot ... `R code` 
\small
```{r ,fig.height=5,fig.width=12}
plot(birth$matage,birth$bweight,las=1,xlab="Mother's Age",
     ylab="Birth weight",col.lab="darkgreen", cex.lab=1.2,
     main="Weight at Birth",col.main="blue",cex.main=1.4,
     col="darkred",type="p", cex=0.9,pch=1)
```


##      


\begin{center}

\textbf{Box plot}

\end{center}

## Box plot
\begin{itemize}
\item [-] Provides a standardized way of displaying the distribution of data. \item []

\item [-] It attempts to provide a visual shape of the data distribution.
\item []
\item [-] This is based on some summary measures: min, $1^{st}$ quartile, median, $3^{rd}$ quartile, and max.   
\item []
\item [-] Range, IQR, Outliers= $3^*IQR$ above $3^{rd}$ or below $1^{st}$ quartiles.   
\end{itemize}

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.75\textwidth,height =0.35\textheight]{box}
\caption{Boxplot}
\end{figure}

[//]: # ## Box plot ...
[//]: # \begin{figure}[htbp!]
[//]: # \centering
[//]: # \includegraphics[width=0.75\textwidth,height =0.85\textheight]{box1}
[//]: # \caption{Boxplot}
[//]: # \label{Boxplot}
[//]: # \end{figure}

## Box plot ...
\begin{figure}[htbp!]
\centering
\includegraphics[width=0.75\textwidth,height =0.55\textheight]{box2}
\end{figure}

[//]: # ## Box plot ...
[//]: # \begin{figure}[htbp!]
[//]: # \centering
[//]: # \includegraphics[width=0.75\textwidth,height =0.85\textheight]{box3}
[//]: # \caption{Boxplot}
[//]: # \label{Boxplot}
[//]: # \end{figure}

[//]: # ## Box plot ...
[//]: #\small
[//]: #```{r,include=FALSE}
[//]: #attach(birth)
[//]: #```
[//]: #```{r,fig.height=5,fig.width=10}
[//]: #par(mfrow=c(1,2))
[//]: #boxplot(matage,las=1,main="Mother's Age");
[//]: #boxplot(gestwks,las=1,main="Gestation")
[//]: #```

## Box plot ...
\small
```{r,include=FALSE}
attach(birth)
```

```{r,fig.height=5,fig.width=9}
boxplot(bweight~sex,las=1,main="Birth weight distribution",
        ylab="birth weight(g)",ylim=c(500,max(bweight)))

```

[//]: # ## Box plot ...
[//]: # \small
[//]: # ```{r,fig.height=5.4,fig.width=6}
[//]: # boxplot(birth[,c("matage","gestwks")],las=1,
[//]: #         names=c("Mother's Age","Gestation"))
[//]: # ```

##

\begin{center}

\textbf{Bar plot}

\end{center}

## Bar plot  

\begin{itemize}

\item [-] Provide a visual presentation of categorical data.
\item []
\item [-] Present grouped data with rectangular bars with lengths proportional to the values that they represent.    
%\item []
%\item [-] uses either horizontal or vertical bars to show comparisons among categories.    
\item []
\item [-] Two types;
  \begin{itemize}
    \item [-.] Grouped - presents bars clustered in groups
    \item [-.] Stacked - shows bars divided into subparts to show cumulative effects.
  \end{itemize}
\end{itemize}

## Bar plot ...
e.g. Cross tabulation of mother's age and birth weight.       

\small
```{r,include=F}
library(xtable)

(Count<-table(birth$lbw,birth$agegrp))
xtable(Count)
Count1<-round(prop.table(Count, 2),2)*100
```
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
 & 20-29 yrs & 30-34 yrs & 35-39 yrs & 40+yrs \\ 
  \hline
Weight $<$2500 &  15(0.16) &  29(0.12) &  32( 0.12) &   4(0.10) \\ 
  Normal 2500+ &  77(0.84) & 222(0.88) & 226(0.88) &  36 (0.90) \\ \hline
   Total    &  92 & 251 & 258 &  40 \\
   \hline \hline
\end{tabular}
\end{table}
## Bar plot ...`Stacked Bar Plot`
\tiny
```{r,fig.height=5,fig.width=10}
barplot(Count,las=1, main="Stacked Bar Plot",border="darkblue",xlab="Age group",ylab="Counts",
        col=c("darkgreen","lightgrey"),legend=rownames(Count))
box();box("figure") 
```

## Bar plot ... `Grouped Bar Plot`
\tiny
```{r,fig.height=4.4,fig.width=8}
barplot(Count, beside = T, main ="Grouped Bar Plot",xlab="Age group",ylab ="Counts"
        , border ="red",yaxt='n',col=c("lightblue","lavender"),ylim=c(0,250),
        legend=rownames(Count),args.legend =list(x ="topright"),space=c(0.05,0.5))
axis(2, at = seq(0, 250, by = 30), las = 1);box()
```


## Saving plot 

Some useful formats.  
\small
```{r, eval=FALSE}
# Start device driver

pdf("mygraph.pdf")  Produces a PDF file   
jpeg() jpeg file  
png()  png file
bmp()  bmp file
win.metafile() windows metafile
postscript() creating PostScript graphics files.   


# terminate device
dev.off()
```
## Saving plot ...
\small
```{r, eval=FALSE}

pdf("barPlot.pdf",width =, height =)

barplot(Count,las=1, main="Stacked Bar Plot",border="darkblue",
      xlab="Age group",ylab="Counts",col=c("darkgreen","lightgrey"),
      legend=rownames(Count))
box()
box("figure")
        
dev.off()

```

## Links 

[Graphical Parameters](http://www.statmethods.net/advgraphs/parameters.html)     
[Introduction to Graphics in r](http://bcb.dfci.harvard.edu/~aedin/courses/Bioconductor/2.Plotting.pdf)
[R Graphs Gallery](http://www.r-graph-gallery.com/all-graphs/)

##

\begin{center}

\textbf{Asante}

\end{center}





